Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-09-15 grove-paper.tex
Fix LaTeX build errors

TODO: PLDI: two column format

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{+}_{4}$" [> "left", > "19"'],
    "$\texttt{*}_{6}$" [> "left", > "19"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
     "$\texttt{*}_{6}$" [> "left", > "19"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
     "$\texttt{+}_{4}$" [> "left", > "19"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
   }
  }
 }
};
\path[draw] (v1) -> (v2);
\path[draw] (v2) -> (v1);
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-User Cycles}
\label{fig:Multi-User Cycles}
\end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
   },
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
     "$\texttt{2}_{4}$" [> "left", > "19"'],
    }
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
     "$\texttt{1}_{4}$" [> "left", > "19"'],
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (-1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
  "$\texttt{2}_{4}$" [> "left", > "19"'],
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
  "$\texttt{1}_{4}$" [> "left", > "19"'],
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Disconnection}
\label{fig:Disconnection}
\end{figure}

@@

%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count
\section{Appendix}
\label{apx:Appendix}

Text of appendix \ldots

@@

>>>>

TODO: fix the following warning which are reported by LaTeX but not Overleaf: Class acmart Warning: A possible image without description on input line 321.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{+}_{4}$" [> "left", > "19"'],
    "$\texttt{*}_{6}$" [> "left", > "19"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Multi-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    "$\texttt{+}_{4}$" [> "left", > "19"'] -> {
     "$\texttt{*}_{6}$" [> "left", > "19"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Multi-User Cycles-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
     "$\texttt{+}_{4}$" [> "left", > "19"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Multi-User Cycles-c}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
   },
   {
    v1/"$\texttt{*}_{6}$" [> "left", > "19"'],
    v2/"$\texttt{+}_{4}$" [> "left", > "19"']
   }
  }
 }
};
\path[draw] (v1) -> (v2);
\path[draw] (v2) -> (v1);
\end{tikzpicture}
\caption{Initial code}
\label{fig:Multi-User Cycles-d}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-User Cycles}
\label{fig:Multi-User Cycles}
\end{figure}

@@

\begin{figure*}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
   },
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
     "$\texttt{2}_{4}$" [> "left", > "19"'],
    }
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
     "$\texttt{1}_{4}$" [> "left", > "19"'],
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-c}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (-1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
  "$\texttt{2}_{4}$" [> "left", > "19"'],
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
  "$\texttt{1}_{4}$" [> "left", > "19"'],
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-d}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Disconnection}
\label{fig:Disconnection}
\end{figure*}

@@

%% Appendixes that will be published  go **BEFORE** the bibliography and count towards the page count

@@

\clearpage % Put these appendixes on separate pages so we can easily remove them from the document.  Also ensure all figures have been placed.

\section{Appendix}
\label{apx:Appendix}

\subsection{Concepts}

\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
  $\mathcal{A}$ & Actions & $\mathcal{S}$ & Edge States  \\
  $\mathcal{C}$ & Cursors & $\mathcal{U}$ & UUIDs        \\
  $\mathcal{E}$ & Edges   & $\mathcal{V}$ & Vertexes     \\
  $\mathcal{G}$ & Graphs  & $\textbf{A}$ & Graph Actions \\
  $\mathcal{I}$ & Indices & $\textbf{E}$ & Editors       \\
  $\mathcal{K}$ & Constructors \\
\end{tabular}

A \emph{UUID} $u \in \mathcal{U}$ is a universally unique identifier.

\subsubsection{The Target Language}

\paragraph{Constructor} ~

A \emph{constructor} $k \in \mathcal{K}$ represents a node in the target language's abstract syntax tree (AST). Let $k_{root}$ be the root AST node.

\paragraph{Index} ~

An \emph{index} $i \in \mathcal{I}$ identifies a position, relative to some parent constructor, of a potential child node in the target language's AST. Let $i_{root}$ be the hyper-position of all rooted constructors.

\subsubsection{The Graph}
\label{sec:concepts:the-graph}

\paragraph{Vertex} ~

A \emph{vertex} $v = (k, u) \in \mathcal{V} = \mathcal{K} \times \mathcal{U}$ is a distinct instance of constructor $k$; that is, one that can be identified by UUID $u$. Let $v_{root} = (k_{root}, u_{root})$ be the root vertex of a graph. The \emph{rooted vertexes} of a graph are the immediate children of its $v_{root}$.

\paragraph{Cursor} ~

A \emph{cursor} $c = (v,i) \in \mathcal{C} = \mathcal{V} \times \mathcal{I}$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$. Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference
to all of the rooted vertexes, of a graph.

\paragraph{Edge} ~

An \emph{edge} $\e = (c, v, u) \in \mathcal{E} = \mathcal{C} \times \mathcal{V} \times \mathcal{U}$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.

\paragraph{Edge State} ~

An \emph{edge state} $s \in \mathcal{S} = \{\bot,+,-\}$ determines whether an edge does not exist ($\bot$), has been created (+), or has been destroyed (-).

\paragraph{Graph Action} ~

A \emph{graph action} $A = (\varepsilon,s,u) \in \mathcal{A} = \mathcal{E} \times \mathcal{S} \times \mathcal{U}$ is an instance of a binding from edge $\varepsilon$ to edge state $s$ that can be identified by UUID $u$. Denote by $[\varepsilon :-> s]_{u}$ a binding from edge $\varepsilon$ to edge state $s$ (identified by UUID $u$).

\paragraph{Graph} ~

A \emph{graph} $G = [\varepsilon :-> s]_{u}^{*} \in \mathcal{G} = \mathcal{A}^{*}$ is a function from edges to edge states. Denote by $G(\varepsilon) = s$ the state $s$ of edge $\varepsilon$ in graph $G$, and by $G A$ the extension of graph $G$ by graph action $A$.

\subsubsection{The Editor}

\paragraph{Move Action} ~

A \emph{move action} $a \in \mathcal{A}^{move} = \{ \text{Left},\text{Right},\text{Up},\text{Down} \} \cup \{ \text{Select} \} \times \mathcal{C}$ re-positions the cursor.

\paragraph{Edit Action} ~

An \emph{edit action} $a \in \mathcal{A}^{edit}= \{ \text{Create} \} \times \mathcal{K} \cup \{ \text{Destroy} \} \cup \{ \text{Restore} \} \times \mathcal{V}$ adds or removes an edge at the cursor.

\paragraph{Communication Action} ~

A \emph{communication action} $a = (\text{Send}, A^{*}, u^{*}) \in \mathcal{A}^{comm} = \{ \text{Send} \} \times \textbf{A}^{*} \times \mathcal{U}^{*}$ applies the graph action sequence $A^{*}$ to the editors identified by UUIDs $u^{*}$.

\paragraph{Action} ~

An \emph{action} $a \in \mathcal{A} = \mathcal{A}^{move} \cup \mathcal{A}^{edit} \cup \mathcal{A}^{comm}$ describes a change to one or more editors.

\paragraph{Editor} ~

An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \textbf{E} = \mathcal{G} \times \mathcal{C} \times \textbf{A}^{*} \times \textbf{A}^{*} \times \mathcal{U}$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.

\subsubsection{The Environment}

An \emph{environment} $(E^{*}, A^{*}) \in \textbf{E}^{*} \times \textbf{A}^{*}$ is a set of communicating editors $E^{*}$ that share a global graph action history $A^{*}$.

\subsection{The Target Language}

\paragraph{Formal Syntax}

$e ::= \lambda p : \tau.e \mid e~e \mid e+e \mid n \mid \_ $

$p ::= x $

$\tau ::= \tau \rightarrow \tau \mid \text{Num}$

\subsection{Sets and Relations}

\paragraph{Edges}

\newcommand{\edges}[1]{TODO:#1}
\newcommand{\vertexes}[1]{TODO:#1}
\newcommand{\liveEdges}[1]{TODO:#1}
\newcommand{\vertexChildren}[1]{TODO:#1}
\newcommand{\cursorChildren}[1]{TODO:#1}
\newcommand{\parents}[1]{TODO:#1}
\newcommand{\dom}[1]{TODO:#1}
\newcommand{\childVertexes}[1]{TODO:#1}
\newcommand{\liveVertexes}[1]{TODO:#1}
\newcommand{\multiparents}[1]{TODO:#1}
\newcommand{\known}[1]{TODO:#1}
\newcommand{\orphans}[1]{TODO:#1}

$\edges(G) = \dom(G)$

$\liveEdges(G) = \{ \e \in \edges(G) | G(\e)=+ \}              \\
  \vertexChildren(G,v) = \{ ((v',i),v'',u) \in \liveEdges(G) | v'=v \} \\
  \cursorChildren(G,c) = \{ (c',v,u) \in \liveEdges(G) | c'=c \}       \\
         \parents(G,v) = \{ (c,v',u) \in \liveEdges(G) | v'=v \}
$

\paragraph{Vertexes} ~

\[
\begin{array}{lr}
\vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
 \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
         \orphans(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || = 0 \}       \\
    \multiparents(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || > 1 \}       \\
        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
    %    \reachable(G) &= \orphans(G) \cup \multiparents(G) \cup \childVertexes(G,v \in \reachable(G)) \\
    % \deleted(G) &= (\orphans(G) \setminus \{ v_{root} \})
    %                \cup (\vertexes(G) \setminus  \\
\end{array}
\]

\paragraph{Graph Actions} ~

$\known(E_1 \cdots E_n) = A_{H1}^{*} \cdots A_{Hn}^{*}
\iff (\forall j=1,\ldots,n)(E_j = (G_j,c_j,A_{Qj}^{*},A_{Hj}^{*},u_j))$
